*Abstract.* The merits of the ``plyr`` package are illustrated.

**1. Introduction.** 

The base R system provides ``lapply()`` and related functions, and the package ``plyr`` provides alternatives that are worth considering.  It will be assumed that readers are familiar with ``lapply()`` and are willing to spend a few moments reading the ``plyr`` documentation, to see why the illustration here will use the ``ldply()`` function.

The test task will be extraction of latitude (and then both latitude and longitude) from the ``section`` dataset in the ``oce`` package.  (Users of that package may be aware that there is a built-in accessor for doing this, so results can easily be checked.)

**2. Methods**

First, load the data

```{r}
library(oce)
data(section)
```

Next, find latitudes using ``lapply``
```{r}
lat <- unlist(lapply(section[["station"]], function(x) x[['latitude']]))
```

Next, find latitudes with ``ldply``
```{r}
library(plyr)
lat <- ldply(section[["station"]], function(x) x[['latitude']])
```

**3. Results**

The reader can check that the results match, although ``ldply()`` returns a data frame, not a vector as in the first method.  Tests of speed
```{r}
library(microbenchmark)
microbenchmark(ldply(section[["station"]], function(x) x[['latitude']])$V1, unlist(lapply(section[["station"]], function(x) x[['latitude']])))
```
indicate such a difference too small to be of much interest, at least in this case.

**4. Discussion**

Since ``ldply()`` returns a data frame, it is more flexible than ``unlist()``, which returns a vector.  For example, the following creates a data frame with columns for lat and lon:
```{r}
latlon <- ldply(section[["station"]], function(x) c(x[['latitude']],x[['longitude']]))
```

A station plot is produced as follows.
```{r}
if (!interactive()) png("plyr.png", unit="in", res=150,
                        width=7, height=7, pointsize=10)
mapPlot(coastlineWorld,
        projection="orthographic", orientation=c(20,-40,0))
mapPoints(latlon$V2, latlon$V1, pch='+', cex=1/2, col='red')
if (!interactive()) dev.off()
```


**5. Conclusions**

The effort of learning how to use the ``plyr`` package is likely to pay off in more flexible code, particularly because of the use of data frames in that package.  On this theme, note that the author of ``plyr`` is developing a similar package called ``dplry``, which centres more closely on data frames and offers many new features; see [http://blog.rstudio.org/2014/01/17/introducing-dplyr/](http://blog.rstudio.org/2014/01/17/introducing-dplyr/) for a blog item introducing ``dplyr``.
